Sequential algorithms to split and merge terabyte-scale 3D images

### 2017-06-07 paper.tex
Wrote remaining algorithms

Algorithms~\ref{algo:naive-slices} and~\ref{algo:naive-blocks} are the
naive merging methods for slices and blocks.
\begin{algorithm}[h]
\caption{Naive merging of slices.}
\label{algo:naive-slices} 
\begin{algorithmic}
  \FOR{each slice}
    \STATE read slice
    \STATE write slice in reconstructed image
  \ENDFOR      
\end{algorithmic}
\end{algorithm}

\begin{algorithm}[h]
\caption{Naive merging of blocks.}
\label{algo:naive-blocks}
\begin{algorithmic}
  \FOR{each block}
    \STATE read block
    \STATE write block in reconstructed image
  \ENDFOR 
\end{algorithmic}
\end{algorithm}

@@

Algo 2: reconstruction from sorted blocks.
* Sort blocks according to (x,y,z)
* For each block in sorted list:
  - read block
  - write block in reconstructed image

@@

@@

We fill memory with as many blocks as possible to write in slices. A
given block is read only once. Slices may be written partially.

Algo 3: reconstruction from blocks using cluster reads and sorted blocks
* Sort blocks according to (x,y,z)
* While there are unread blocks:
  * While there is memory available:
    - read next block in sorted list
  * (Partially) reconstruct slices from blocks (in memory).
  * For each partially reconstructed slice:
    - Write in reconstructed image

@@

We fill memory with as many voxels as possible to write groups of
entire slices. A given block may be read multiple times.

Algo 4: reconstruction from blocks using multiple reads
* While there are empty slices in the reconstructed image:
  * Determine $n$ so that $n$ slices fit in memory
  * Reconstruct next slices $s_1$, ...,  $s_n$:
     - list the blocks contributing to $s_1$, $s_n$
     - for each block in the list, read the voxels of interest
     - reconstruct $s_1$, ..., $s_n$ in memory
     - write $s_1$, ..., $s_n$

>>>>

Algorithms~\ref{algo:naive-slices} and~\ref{algo:naive-blocks} are the
naive merging methods for slices and blocks.
\begin{algorithm}[h]
\caption{Naive merging from slices.}
\label{algo:naive-slices} 
\begin{algorithmic}
  \FOR{each slice}
    \STATE read slice
    \STATE write slice in reconstructed image
  \ENDFOR      
\end{algorithmic}
\end{algorithm}

\begin{algorithm}[h]
\caption{Naive merging from blocks.}
\label{algo:naive-blocks}
\begin{algorithmic}
  \FOR{each block}
    \STATE read block
    \STATE write block in reconstructed image
  \ENDFOR 
\end{algorithmic}
\end{algorithm}

@@

\begin{algorithm}[h]
  \caption{Merging from sorted blocks.}
  \label{algo:sorted-blocks}
  \begin{algorithmic}
    \STATE list = sort blocks according to (x,y,z)
    \FOR{each block in list}
      \STATE read block
      \STATE write block in reconstructed image
    \ENDFOR
  \end{algorithmic}
  \end{algorithm}

@@

\begin{algorithm}[h]
  \caption{Buffered merging from slices}
  \label{algo:buffered-slices}
  \begin{algorithmic}
    \STATE list = sort slices by increasing x values
    \STATE initialize buffer
    \STATE index = 0
    \WHILE{index\textless n}
      \WHILE{sizeof(buffer)\textless m}
        \STATE slice = list[index]
        \STATE read slice and append it to buffer
        \STATE index += 1
       \ENDWHILE
       \STATE write buffer in reconstructed image
       \STATE clear buffer
    \ENDWHILE
  \end{algorithmic}
\end{algorithm}

@@

See Algorithm~\ref{algo:cluster-reads}. As many blocks as possible are
loaded in memory, concatenated in a buffer and written as sequentially
as possible from the buffer. Although the Algorithm looks very similar
to Algorithm~\ref{algo:buffered-slices}, it should be noted that in
Algorithm~\ref{algo:cluster-reads} the buffer should actually be a
more complex data structure capable of storing multiple disjoint
sequences of contiguous bytes without having to allocate memory to
store the ``holes'' between such sequences \todo{a figure showing the
  buffer and how it fillus up would help}.  \note{More on
  defragmentation might be needed here.}
\begin{algorithm}[h]
  \caption{Buffered merging from blocks using cluster reads}
  \label{algo:cluster-reads}
  \begin{algorithmic}
    \STATE list = sort blocks by increasing (x,y,z) values
    \STATE initialize buffer
    \STATE index = 0
    \WHILE{index\textless n}
      \WHILE{sizeof(buffer)\textless m}
        \STATE block = list[index]
        \STATE read block and append it to buffer
        \STATE index += 1
        \ENDWHILE
       \STATE write buffer in reconstructed image
       \STATE clear buffer
    \ENDWHILE
  \end{algorithmic}
\end{algorithm}

@@

In multiple reads, blocks are read partially to ensure that the memory
buffer only contains contiguous bytes, at any time. That is, a given
block will be read multiple times.

\begin{algorithm}[h]
  \caption{Buffered merging from blocks using multiple reads
    \todo{improve notations and algo}.}
  \label{algo:multiple-reads}
  \begin{algorithmic}
    \STATE list = sort blocks by increasing (x,y,z) values
    \STATE initialize buffer
    \STATE pointer = 0, max\_index=(m-1)
    \WHILE{max\_index \textless sizeof(reconstructed\_image)}
    \WHILE{pointer \textless max\_index}
        \STATE block = read block that contains voxel index\_to\_voxel(pointer)
        \STATE insert in buffer all voxels (x,y,z) of block where voxel\_to\_index(x,y,z) \textless max\_index
        \STATE update pointer to point to the next unassigned element in buffer
        \ENDWHILE
       \STATE write buffer in reconstructed image
       \STATE clear buffer
       \STATE pointer = max\_index + 1
       \STATE max\_index += m
    \ENDWHILE
  \end{algorithmic}
\end{algorithm}
